ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-08 rebuttal.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer \#2:}

\begin{reviewer}
    %
    My main concern with the creation of a publically available database, formatted with ChemKED, is
    that it seems that the creation of the files has to be done manually, i.e. susceptible for
    errors. This is especially true if the experiment comprises a big set of datapoints and/or
    hundreds of measured species. Furthermore, the YAML data serialization format is indeed
    human-readable, but it does not provide an easy summary/overview of the experimental data, such
    as an excel or CSV file would give, and therefore, people might be not willing to put the extra
    effort.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer, insofar as manual creation of the files may lead to errors in the
    encoded data. However, this will be the case wherever manual processing is required, regardless
    of the database format, and could even be the case for tables or figures published in journal
    articles. Thus, we do not feel that manual file creation is an inherent disadvantage of our
    format. Moreover, as we discuss below, we are working with experimental teams to integrate the
    creation of ChemKED files into their workflows to try to mitigate some of these problems.

    On the second point, we agree with the reviewer that CSV or Excel files may give an adequate
    summary or overview of the data, but only if the files are opened with some appropriate
    software, e.g., Microsoft Excel (particularly for Excel-formatted files). Thus, in this sense,
    CSV or Excel files do not offer any advantage over ChemKED files (or any other database files)
    that must also be processed with an appropriate software package (PyKED, in the case of ChemKED
    files) so that a summary can be produced. Furthermore, in the specific case of PyKED/ChemKED, it
    is already possible to produce an Excel or CSV file (using the Pandas DataFrame integration) so
    that one may view such a summary if desired. See also the response to a similar comment below.
    %
\end{response}

### 2017-10-08 rebuttal.tex

\subsection*{Overview of ChemKED format}

\begin{reviewer}
    %
    9.~A reference should not be a required mapping of an experimental dataset. If the intention is
    to make the community to adopt the ChemKED file structure, it might be easier to do so if
    experimentalists can use the structure internally as well, i.e. to read, store, and share
    non-published internal data.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
    users who wish to use ChemKED files in this way include an incomplete reference field. We have
    modified the ChemKED schema in response to this comment so that the only required subfields of
    the reference field are the authors and the year.

    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor,
    which turns off all validation of the ChemKED file when it is loaded with PyKED. Work is ongoing
    to allow more granular selection of fields to be validated, and discussion is contained in a
    public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
    %
\end{response}

### 2017-10-08 rebuttal.tex

\begin{reviewer}
    %
    11.~Why do both \texttt{atomic-composition} and \texttt{elemental-composition} exist as
    synonyms? Wouldn't it be clearer if only one of the two can be used as a valid input?
    %
\end{reviewer}

\begin{response}
    We agree with the reviewer on this point, and have removed \texttt{elemental-composition}
    as an option for clarity. The text on \textbf{page 6} has been updated to reflect this.
\end{response}

\begin{reviewer}
    %
    12.~Is it possible to create for example two \texttt{common-properties} \texttt{composition}
    blocks A and B and refer to them in the \texttt{DataPoint} keys?
    %
\end{reviewer}

\begin{response}
    %
    At present, this is not possible; only a single \verb|composition| block is allowed in the
    \verb|common-properties| section. However, we are considering modifying the behavior of the
    \verb|common-properties| section, and discussion is ongoing in
    \url{https://github.com/pr-omethe-us/PyKED/issues/67}.
    %
\end{response}

@@

\begin{reviewer}
    %
    13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED
    files, on GitHub. Is this the most efficient way for making a publically available database? One
    issue with this is the fact that adding files is done via pull requests, which have to be
    verified and approved by the owners of the repository. Therefore a certain delay is inevitable.
    %
\end{reviewer}

\begin{response}
    %
    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some
    delay, however we see the human intervention step as necessary and important. In much the same
    way that review of manuscripts provides a certain surety of the minimal quality of the article,
    we see review of data being added to the database as a minimal quality check; for instance, does
    the reference exist, has the data already been added to the database, are the files styled
    consistently, etc. Furthermore, such human review can help to catch some errors that may be
    introduced by manual creation of the files, as pointed out by the reviewer previously.

    Moreover, all the existing databases have a similar requirement (whether explicitly stated or
    not), where users submit data files for inclusion that are approved by the database owners after
    some review. In this sense, we actually see the Pull Request mechanism of GitHub as an
    advantage, since all reviews can be done publicly and some of the review procedures (style
    checks, reference validity, etc.) may be done automatically by a script associated with the
    GitHub repository. In this way, the entire community can contribute to improving the quality of
    the database.

    Lastly, while we hope other researchers will contribute to this database (such that it becomes a
    true community effort), there is no requirement for others to use our database to use ChemKED
    files or PyKED. Others are free to use the tools internally, or download\slash fork our database
    and build their own. Moreover, as mentioned in the manuscript, we have included the ability to
    generate ReSpecTh-compliant files from ChemKED files, so that the community can move data to
    alternative formats easily and so there is a reduced chance of a negative ``lock-in'' effect.
    %
\end{response}

### 2017-10-08 pyked.tex
Update open definition to include open knowledge

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/})
has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML can represent detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be much more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them. Furthermore, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

%discuss open source definition here
(When we describe a software package or project as ``closed,'' we are generally
contrasting it with the open source definition~\cite{opensource}. In brief, to be
``open,'' a project must provide free and unrestricted access to source code and\slash
or data, allow free redistribution, permit derived works, and not discriminate against
any people or areas of work, along with a few other criteria.)

>>>>

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/})
has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem.

Third, XML is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML can represent detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be much more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them.

Finally, the closed and opaque natures of the PrIMe standard and associated Data Warehouse limit
contributions and community development of the PrIMe standard and compatible data files and tools.
In this sense, we mean ``closed'' as a contrast to open source, open science, and open data, as
defined by the Open Knowledge Institute~\cite{openknowledge} or the Open Source
Initiative~\cite{opensource}. Although most commonly applied to software, the distinction between
open and closed is becoming increasingly important in science and research. In brief, the Open
Knowledge Institute defines open knowledge as ``free to access, use, modify, and share... — subject,
at most, to measures that preserve provenance and openness''~\cite{openknowledge}, while the Open
Source Institute focuses on software and declares that a project must provide free and unrestricted
access to the source code, allow free redistribution, permit derived works, and not discriminate
against any people or areas of work, along with a few other criteria~\cite{opensource}.
